---
title: 'Emergency 911 Calls - Montgomery County, PA'
date: "October 15, 2016"
output: html_document
---

***

### About 911:

* Created by Congress in 2004 as the 911 Implementation and Coordination Office (ICO), the National 911 Program is housed within the National Highway Traffic Safety Administration at the U.S. Department of Transportation and is a joint program with the National Telecommunication and Information Administration in the Department of Commerce.

For more information : <http://www.911.gov/about_national_911program.html>

### About Montgomery County

* Montgomery County, locally also referred to as Montco, is a county located in the Commonwealth of Pennsylvania. As of the 2010 census, the population was 799,874, making it the third-most populous county in Pennsylvania, after Philadelphia and Allegheny Counties. The county seat is Norristown. Montgomery County is very diverse, ranging from farms and open land in Upper Hanover to densely populated rowhouse streets in Cheltenham.


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(ggmap)
library(viridis)
library(plotly)
```

### Sneak peak into the dataset

```{r echo=FALSE, warning=FALSE}

mydata <- read.csv("../input/911.csv")

dim(mydata)

str(mydata)

summary(mydata)

```


* Zip should be a factor
* time stamp should be of type date 
* Can remove dummy variable if it is always 1

### Cleaning the dataset

* Changing the zip class
* Removing the dummy variable
* Changing the class of timestamp

```{r echo=FALSE, warning=FALSE}

mydata$zip <- factor(mydata$zip)

mydata <- mydata[,-9]

mydata$timeStamp <- as.POSIXct(mydata$timeStamp)

mydata$Date <- as.Date(mydata$timeStamp)

mydata <- separate(mydata, col = title, into = c("Type", "SubType"), sep = ":")

mydata$SubType <- gsub(" -", "", mydata$SubType)

```


### Creating new variables

* Creating type and subtype based on the title column
* Creating the Year and Month from the timestemp column

```{r echo=FALSE, warning=FALSE}

mydata$Year <- year(mydata$timeStamp)
mydata$Month <- month(mydata$timeStamp)
mydata$Day <- day(mydata$timeStamp)
mydata$Hour <- hour(mydata$timeStamp)
mydata$Weekday <- weekdays(mydata$timeStamp)

mydata$Year <- factor(mydata$Year)
mydata$Month <- factor(mydata$Month)
mydata$Day <- factor(mydata$Day)
mydata$Hour <- factor(mydata$Hour)
mydata$Weekday <- factor(mydata$Weekday)

mydata <- mydata[,-7]

```

### Number of calls over the period of time and by type?

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

by_date <- mydata %>% group_by(Date) %>% summarise(Total = n())

head(by_date)

ggplot(by_date, aes(Date, Total)) + geom_line(color = "blue", size = 1)

by_date_type <- mydata %>% group_by(Date, Type) %>% summarise(Total = n())

by_date_type$Type <- factor(by_date_type$Type)

ggplot(by_date_type, aes(Date, Total)) + geom_line( aes(color = Type), size = 0.6)

ggplot(by_date_type, aes(Date, Total)) + geom_line( aes(color = Type), size = 0.6) + facet_wrap(~Type) + theme(legend.position="none")
```

* Only few times Traffic calls have exceeded the EMS calls
* Only on Jan 23 2016 number of traffic calls has almost gone up by 3 to 4 times.

### How many calls hourly, monthly and yearly?

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

table(mydata$Year)

table(mydata[mydata$Year==2016,]$Month)

table(mydata[mydata$Year==2016,]$Month)

ggplot(mydata[mydata$Year==2016,], aes(Month, fill = Month)) + geom_bar() + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 - Monthly ")

ggplot(mydata, aes(Hour, fill = Hour)) + geom_bar() + theme(legend.position = "none") + ggtitle("911 Emergency Calls - Hourly")

ggplot(mydata, aes(Weekday, fill = Weekday)) + geom_bar() + theme(legend.position = "none") + ggtitle("911 Emergency Calls - Weekday")

```

* First table gave information about calls year wise but when looked at the data more closer found that 2015 records have only calls of December.
* January has highest number of calls so far this year with July being the second 
* Majority of the calls are during day time.

### How many calls based on type?

```{r echo=FALSE, warning=FALSE, fig.width=8, fig.height=5}

table(mydata$Type)

ggplot(mydata, aes(Type, fill = Type)) + geom_bar() + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 - Type")

prop.table(table(mydata$Type))

ggplot(as.data.frame(prop.table(table(mydata$Type))), aes(Var1, Freq, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 - Type Percentage") + xlab("Type") + ylab("Percentage of Calls")


```

* Nearly 50% of the calls are of Type EMS
* Traffic related come second.

### How many calls from each subtypes in overall dataset and also subtypes within each type?

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=5.5}

top_subtypes <- as.data.frame(table(mydata$SubType))
top_subtypes <- top_subtypes[order(-top_subtypes$Freq),]
top10_subtypes <- top_subtypes[1:10,]
top10_subtypes$Perc <- top10_subtypes$Freq/sum(top_subtypes$Freq) * 100
top10_subtypes

ggplot(top10_subtypes, aes(reorder(Var1, Freq), Freq, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls") + xlab("Subtype") + ylab("Number of 911 Calls")  + coord_flip()

ggplot(top10_subtypes, aes(reorder(Var1, Perc), Perc, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls") + xlab("Subtype") + ylab("Percentage of 911 Calls")  + coord_flip()


gettop10subtypes <- function(type) {
  mytype <- subset(mydata, Type == type)
  mytype$SubType <- factor(mytype$SubType)
  mytype_subtypes <- as.data.frame(table(mytype$SubType))
  mytype_subtypes <- mytype_subtypes[order(-mytype_subtypes$Freq),]
  top10_types_substype <- mytype_subtypes[1:10,]
  top10_types_substype$Perc <- top10_types_substype$Freq/sum(mytype_subtypes$Freq) * 100
  return(top10_types_substype)
}

gettop10subtypes("EMS")

ggplot(gettop10subtypes("EMS"), aes(reorder(Var1, Freq), Freq, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016") + xlab("Subtype") + ylab("Number of EMS Calls")  + coord_flip()

ggplot(gettop10subtypes("EMS"), aes(reorder(Var1, Perc), Perc, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016") + xlab("Subtype") + ylab("Percentage of EMS Calls")  + coord_flip()


gettop10subtypes("Fire")

ggplot(gettop10subtypes("Fire"), aes(reorder(Var1, Freq), Freq, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 ") + xlab("Subtype") + ylab("Number of Fire Calls")  + coord_flip()

ggplot(gettop10subtypes("Fire"), aes(reorder(Var1, Perc), Perc, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 ") + xlab("Subtype") + ylab("Percentage of Fire Calls")  + coord_flip()

gettop10subtypes("Traffic")[1:7,]

ggplot(gettop10subtypes("Traffic")[1:7,], aes(reorder(Var1, Freq), Freq, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016") + xlab("Subtype") + ylab("Number of Traffic Calls")  + coord_flip()

ggplot(gettop10subtypes("Traffic")[1:7,], aes(reorder(Var1, Perc), Perc, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 ") + xlab("Subtype") + ylab("Percentage of Traffic Calls")  + coord_flip()

```

* Nearly 29% of all the 911 calls are for Vehicle accident. 
* Nearly 8% are for Disabled vehicle which comes second in overall calls.
* EMS alone has 72 subtypes
* Respiratory and cardiac emergency are the top subtypes among EMS
* Out of all the calls for EMS, Respiratory and Cardiac contribute to 10% of the calls.
* Fall Victim coming third
* Fire alarm calls contribute 37% of the Fire related calls
* Vehicle accidents contribute 11% of Fire calls
* Vehicle accidents contribute 65% of the Traffic calls with disabled vehicles coming second with 21%

### Which zip codes have the highest number of calls?

```{r echo=FALSE, warning=FALSE,fig.width=9, fig.height=5.5}

top_zip <- as.data.frame(table(mydata$zip))
top_zip <- top_zip[order(-top_zip$Freq),]
top10_zip <- top_zip[1:10,]

names(top10_zip) <- c("Zip", "Total")
top10_zip$Perc <- top10_zip$Total/sum(top_zip$Freq) * 100
top10_zip$Zip <- factor(top10_zip$Zip)

top10_zip

ggplot(top10_zip, aes(reorder(Zip, -Total), Total, fill = Zip)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 - Top 10 Zip ") + xlab("Zip codes with most number of calls")

ggplot(top10_zip, aes(reorder(Zip, -Perc), Perc, fill = Zip)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 - Top 10 Zip ") + xlab("Zip codes  Percentage of calls")

```

* Around 45% of the calls are from the top 20 zip codes
* It will be interesting to know the population of all the zip codes to find if there is any relationship between population and count of calls.


### Which township contributes to majority of the calls?


```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

length(unique(mydata$twp))

top_twp <- as.data.frame(table(mydata$twp))
top_twp <- top_twp[order(-top_twp$Freq),]
top10_twp <- top_twp[1:10,]

names(top10_twp) <- c("Twp", "Total")
top10_twp$Perc <- top10_twp$Total/sum(top_twp$Freq) * 100
top10_twp$Twp <- factor(top10_twp$Twp)

top10_twp

ggplot(top10_twp, aes(reorder(Twp, -Total), Total, fill = Twp)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 - Top 10 Townships ") + xlab("Townships with most number of calls")  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5)) 

ggplot(top10_twp, aes(reorder(Twp, -Perc), Perc, fill = Twp)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("911 Emergency Calls for 2016 - Top 10 Townships ") + xlab("Townships Percentage of calls")  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5)) 

```

* Out of 69 townships, top 10 contribute around 48% of the calls
* It would be interesting to know the developmental ratings of these townships for getting more insights


### HeatMaps

* Generating heatmaps of calls by Month, day and hour.

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

## Get the count of the calls by Month and Hour
day_hour <- mydata[mydata$Year == "2016", c("Day", "Hour")] %>% group_by(Day, Hour) %>% summarise(Count = n())
day_hour <- as.data.frame(day_hour)

## Change the type of the variables
day_hour$Day <- as.factor(day_hour$Day)
day_hour$Hour <- as.factor(day_hour$Hour)

## Building heatmap using ggplot2
ggplot(day_hour, aes(Day, Hour, fill = Count)) + geom_tile(color = "white", size = 0.1) + scale_fill_viridis(name="#Calls") + coord_equal() + labs(title="911 Calls by Day and Hour") 


## Get the count of the calls by Month and Day
month_day <- mydata[mydata$Year == "2016", c("Month", "Day")] %>% group_by(Month, Day) %>% summarise(Count = n())
month_day <- as.data.frame(month_day)

## Change the type of the variables
month_day$Month <- as.factor(month_day$Month)
month_day$Day <- as.factor(month_day$Day)

## Change the levels of the Month 
levels(month_day$Month) <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

## Building heatmap using ggplot2
ggplot(month_day, aes(Day, Month, fill = Count)) + geom_tile(color = "white", size = 0.1) + scale_fill_viridis(name="#Calls") + coord_equal() + labs(title="911 Calls by Month and Day") 

```


* Majority of the calls are during the day time.
* There are days when the calls were huge in number specially Jan 23.
* When you look closer at what hour of Jan 23 you find that around 4 and 5 in the evening there were huge number of calls.


### How does the type vary among top 10 township?

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

ggplot(mydata[mydata$twp %in% top10_twp$Twp, ], aes(twp, fill = Type )) + geom_bar(position = "dodge")  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5))  + xlab("Township") + ggtitle("Type of calls among Top 10 Townships")

ggplot(mydata[mydata$twp %in% top10_twp$Twp, ], aes(twp, fill = Type)) + geom_bar()  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5))  + xlab("Township") + ggtitle("Type of calls among Top 10 Townships") + facet_wrap(~Type) + theme(legend.position="none")

```

* Lower and Upper Merion have highest number of Traffic calls
* Norristown has highest number of EMS calls
* Lower Merion also has highest number of Fire calls.
* Is there a infrastructure problem in Lower and Upper Merion? Or Is the traffic calls due to reckless driving by large number of people in those areas? 


### How does the Top 5 subtypes vary among top 10 twp?

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

top3 <- top10_subtypes[1:5,]$Var1

sample <- mydata[mydata$twp %in% top10_twp$Twp,]
sample <- sample[sample$SubType %in% top3, ]
dim(sample)

ggplot(sample, aes(twp, fill = SubType )) + geom_bar(position = "dodge")  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5))  + xlab("Township") + ggtitle("Subtype distribution among Top 10 Townships")

```

* Vehicle accidents are huge among all the townships which is very concerning given the number of lives at risk during each accident

### Vehicle accidents by hour, day and type

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

vehicle <- mydata[mydata$SubType ==" VEHICLE ACCIDENT", ]

table(vehicle$Type)

ggplot(vehicle, aes(Hour, fill = Type)) + geom_bar(position = "dodge") + ggtitle("Vehicle Accidents by Hours")

day_hour <- vehicle[vehicle$Year == "2016", c("Day", "Hour")] %>% group_by(Day, Hour) %>% dplyr::summarise(Count = n())
day_hour <- as.data.frame(day_hour)

day_hour$Day <- as.factor(day_hour$Day)
day_hour$Hour <- as.factor(day_hour$Hour)

ggplot(day_hour, aes(Day, Hour, fill = Count)) + geom_tile(color = "white", size = 0.1) + scale_fill_viridis(name="#Calls") + coord_equal() + labs(title="Vehicle Accident Calls by Day and Hour - 2016")

```

* Majority of the Vehicle Accident calls are during the evening peak hours.
* People should be educated about the rush hour accidents and make them drive with more caution


### CARDIAC EMERGENCY by hour and day

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

cardiac <- mydata[mydata$SubType ==" CARDIAC EMERGENCY", ]

day_hour <- cardiac[cardiac$Year == "2016", c("Day", "Hour")] %>% group_by(Day, Hour) %>% dplyr::summarise(Count = n())
day_hour <- as.data.frame(day_hour)

day_hour$Day <- as.factor(day_hour$Day)
day_hour$Hour <- as.factor(day_hour$Hour)

ggplot(day_hour, aes(Day, Hour, fill = Count)) + geom_tile(color = "white", size = 0.1) + scale_fill_viridis(name="#Calls") + coord_equal() + labs(title="CARDIAC EMERGENCY Calls by Day and Hour - 2016") 

```


* Majority of the Cardiac emergencies are during day time mostly middle part of the day.
* Is it because people are more stressed out during day time?
