---
title: 'Analyzing a Pokémon data set'
author: Xavier Vivancos García 
date: '2018-03-28'
output: 
  html_document:
    number_sections: yes
    toc: yes
    theme: cerulean
---

******
# Loading data 
******

```{r message=FALSE}
# Load libraries
library(tidyverse)
library(gridExtra)

# Read the Pokémon stats
pokemon <- read.csv("../input/Pokemon.csv")

# View first 6 rows
head(pokemon)

# Express the Generation column as a factor
pokemon$Generation <- as.factor(pokemon$Generation)

# Structure of Pokémon data set
str(pokemon)
```

We see that there are duplicate numbers in the `X.` column. Those rows correspond to Mega Evolutions or different versions of the same Pokémon. 

******
# Data Analysis 
******

******
## Types
******

First let's see the different types of Pokémon. 

```{r}
# Primary types
levels(pokemon$Type.1)

# Secondary types
levels(pokemon$Type.2)
```

Is this really necessary? Come on dude, you should already know them! Notice that we have one more level in the secondary types. Why? Because a Pokémon may have either one or two types. For the Pokémon with only one type we have the empty level in the `Type.2` column. 

```{r}
# How many Pokémon of each primary type are there?
ggplot(pokemon, aes(x=fct_infreq(Type.1))) + 
  geom_bar(fill="#99CCFF", colour="black") +
  labs(x="Type 1", y="Frequency") +
  theme(axis.text.x=element_text(angle=45, hjust=1))

# How many Pokémon of each secondary type are there?
ggplot(pokemon, aes(x=fct_infreq(Type.2))) +
  geom_bar(fill="#99CCFF", colour="black") +
  labs(x="Type 2", y="Frequency") +
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

As we see, most Pokémon have only one type. Can we calculate the number of Pokémon for each type combination? Of course! We can create a contingency table containing the counts of each combination of the levels of the variables involved (`Type.1` and `Type.2`). 

```{r}
# Counts of each type combination
types <- pokemon %>%
  group_by(Type.1, Type.2) %>%
  summarise(count=n()) 

# Contingency table of Type.1 and Type.2
ggplot(types, aes(x=Type.1,y=Type.2)) +
  geom_tile(aes(fill=count)) +
  geom_text(aes(label=count)) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  labs(x="Type 1", y="Type 2") +   
  scale_fill_gradient(low="white", high="red") 
```

The last row correspond to Pokémon with one type. For example, there are 27 pure Electric-type Pokémon. 

******
## Stats 
******

Let's start with some simple histograms to represent the distribution of the following stats: HP, Attack, Defense, Special Attack, Special Defense, and Speed.

```{r}
# Histogram of HP
p1 <- ggplot(pokemon, aes(x=HP)) +
  geom_histogram(binwidth=4, fill="palegreen", colour="black") + 
  labs(x="HP", y="Frequency") 

# Histogram of Attack
p2 <- ggplot(pokemon, aes(x=Attack)) +
  geom_histogram(binwidth=4, fill="palegreen", colour="black") + 
  labs(x="Attack", y="Frequency") 

# Histogram of Defense
p3 <- ggplot(pokemon, aes(x=Defense)) +
  geom_histogram(binwidth=4, fill="palegreen", colour="black") + 
  labs(x="Defense", y="Frequency") 

# Histogram of Special Attack
p4 <- ggplot(pokemon, aes(x=Sp..Atk)) +
  geom_histogram(binwidth=4, fill="palegreen", colour="black") + 
  labs(x="Special Attack", y="Frequency") 

# Histogram of Special Defense
p5 <- ggplot(pokemon, aes(x=Sp..Def)) +
  geom_histogram(binwidth=4, fill="palegreen", colour="black") + 
  labs(x="Special Defense", y="Frequency") 

# Histogram of Speed
p6 <- ggplot(pokemon, aes(x=Speed)) +
  geom_histogram(binwidth=4, fill="palegreen", colour="black") + 
  labs(x="Speed", y="Frequency") 

# Histogram of Total 
p7 <- ggplot(pokemon, aes(x=Total)) +
  geom_histogram(binwidth=4, fill="palegreen", colour="black") + 
  labs(x="Total", y="Frequency") 

# Subplot with all the histograms 
grid.arrange(p1, p2, p3, p4, p5, p6, p7,
             layout_matrix=cbind(c(1,4,7), c(2,5,7), c(3,6,7)))
```

Are the Legendary Pokémon much better than the normal ones? We can compare the density plots to see the difference between their stats.  

```{r}
# Density plot of HP
p8 <- ggplot(pokemon, aes(x=HP, fill=Legendary)) +
  geom_density(alpha=0.5) +
  labs(x="HP", y="Density") +
  theme(legend.position="none")

# Density plot of Attack
p9 <- ggplot(pokemon, aes(x=Attack, fill=Legendary)) +
  geom_density(alpha=0.5) +
  labs(x="Attack", y="Density") +
  theme(legend.position="none")

# Density plot of Defense
p10 <- ggplot(pokemon, aes(x=Defense, fill=Legendary)) +
  geom_density(alpha=0.5) +
  labs(x="Defense", y="Density") +
  theme(legend.position="none")

# Density plot of Special Attack
p11 <- ggplot(pokemon, aes(x=Sp..Atk, fill=Legendary)) +
  geom_density(alpha=0.5) +
  labs(x="Special Attack", y="Density") +
  theme(legend.position="none")

# Density plot of Special Defense
p12 <- ggplot(pokemon, aes(x=Sp..Def, fill=Legendary)) +
  geom_density(alpha=0.5) +
  labs(x="Special Defense", y="Density") +
  theme(legend.position="none")

# Density plot of Speed
p13 <- ggplot(pokemon, aes(x=Speed, fill=Legendary)) +
  geom_density(alpha=0.5) +
  labs(x="Speed", y ="Density") +
  theme(legend.position="none")

# Density plot of Total
p14 <- ggplot(pokemon, aes(x=Total, fill=Legendary)) +
  geom_density(alpha=0.5) +
  labs(x="Total", y="Density") 

# Subplot with all the density plots 
grid.arrange(p8, p9, p10, p11, p12, p13, p14,
             layout_matrix=cbind(c(1,4,7), c(2,5,7), c(3,6,7)))
```

It is clear that Legendary Pokémon have greater statistics. What about the boxplots? It may be interesting to see the statistical distribution according to the Pokémon's type. 

```{r}
# Boxplot of HP
ggplot(pokemon %>% group_by(Type.1) %>% mutate(med=median(HP)),
       aes(x=reorder(Type.1, HP, FUN=median), y=HP)) +
  geom_boxplot(aes(fill=med)) +
  scale_fill_gradient(low="paleturquoise", high="paleturquoise4") +
  coord_flip() +
  labs(x="Type 1") +
  theme(legend.position="none")

# Boxplot of Attack
ggplot(pokemon %>% group_by(Type.1) %>% mutate(med=median(Attack)),
       aes(x=reorder(Type.1, Attack, FUN=median), y=Attack)) +
  geom_boxplot(aes(fill=med)) +
  scale_fill_gradient(low="paleturquoise", high="paleturquoise4") +
  coord_flip() +
  labs(x="Type 1") +
  theme(legend.position="none")

# Boxplot of Defense
ggplot(pokemon %>% group_by(Type.1) %>% mutate(med=median(Defense)),
       aes(x=reorder(Type.1, Defense, FUN=median), y=Defense)) +
  geom_boxplot(aes(fill=med)) +
  scale_fill_gradient(low="paleturquoise", high="paleturquoise4") +
  coord_flip() +
  labs(x="Type 1") +
  theme(legend.position="none")

# Boxplot of Special Attack
ggplot(pokemon %>% group_by(Type.1) %>% mutate(med=median(Sp..Atk)),
       aes(x=reorder(Type.1, Sp..Atk, FUN=median), y=Sp..Atk)) +
  geom_boxplot(aes(fill=med)) +
  scale_fill_gradient(low="paleturquoise", high="paleturquoise4") +
  coord_flip() +
  labs(x="Type 1") +
  theme(legend.position="none")

# Boxplot of Special Defense
ggplot(pokemon %>% group_by(Type.1) %>% mutate(med=median(Sp..Def)),
       aes(x=reorder(Type.1, Sp..Def, FUN=median), y=Sp..Def)) +
  geom_boxplot(aes(fill=med)) +
  scale_fill_gradient(low="paleturquoise", high="paleturquoise4") +
  coord_flip() +
  labs(x="Type 1") +
  theme(legend.position="none")

# Boxplot of Speed
ggplot(pokemon %>% group_by(Type.1) %>% mutate(med=median(Speed)),
       aes(x=reorder(Type.1, Speed, FUN=median), y=Speed)) +
  geom_boxplot(aes(fill=med)) +
  scale_fill_gradient(low="paleturquoise", high="paleturquoise4") +
  coord_flip() +
  labs(x="Type 1") +
  theme(legend.position="none")

# Boxplot of Total
ggplot(pokemon %>% group_by(Type.1) %>% mutate(med=median(Total)),
       aes(x=reorder(Type.1, Total, FUN=median), y=Total)) +
  geom_boxplot(aes(fill=med)) +
  scale_fill_gradient(low="paleturquoise", high="paleturquoise4") +
  coord_flip() +
  labs(x="Type 1") +
  theme(legend.position="none")
```

The outliers in the boxplots correspond to Pokémon (normally Legendary or Mega Evolutions) with impressive stats numerically distants from the rest of the data. For example, Regirock is an outlier in the defense boxplot (200 points in defense!). 

```{r}
# Density plot of Total
 ggplot(pokemon, aes(x=Total)) +
   geom_density(alpha=0.5, fill="#56B4E9") +
   facet_wrap(~Type.1) + 
   labs(x="Total", y="Density") +
   theme(axis.text.x=element_blank(), 
         axis.ticks.x=element_blank(),
         legend.position="none")
```

Analyzing the boxplots and the density plots, we can say that the Dragon-type Pokémon are statistically superior from the rest. Another question to answer: In which generation the stats are better?

```{r}
# Average Total for each generation 
pokemon %>%
  group_by(Generation) %>%
  summarize(Total=mean(Total)) %>%
  ggplot(aes(x=Generation, y=Total, group=1)) +
    geom_line(colour="red") +
    geom_point() +
    labs(y="Average Total")
```

The fourth generation has the best average stats, probably because it is the generation with more Legendary Pokémon. What happens if we split up the Total stat?

```{r}
# Average stats for each generation 
pokemon %>%
  group_by(Generation) %>%
  summarize(HP=mean(HP),
            Attack=mean(Attack),
            Defense=mean(Defense),
            Sp..Atk=mean(Sp..Atk),
            Sp..Def=mean(Sp..Def),
            Speed=mean(Speed)) %>%
  gather(Stats, value, 2:7) %>%
  ggplot(aes(x=Generation, y=value, group=1)) +
    geom_line(colour="red") +
    geom_point() +
    facet_wrap(~Stats) +
    labs(y="Average Stats")
```

To end this section we are going to reveal the Top 10 Best Pokémon for each stat. Are you nervous? Let's begin!

```{r}
# Top 10 HP Pokémon 
pokemon %>%
  select(Name, HP) %>%
  arrange(desc(HP)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(Name, HP), y=HP)) +
  geom_bar(stat="identity", fill="lightcyan", colour="black") +
  coord_flip() +
  labs(x="Name", title="Top 10 HP Pokémon")

# Top 10 Attack Pokémon 
pokemon %>%
  select(Name, Attack) %>%
  arrange(desc(Attack)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(Name, Attack), y=Attack)) +
  geom_bar(stat="identity", fill="lightcyan", colour="black") +
  coord_flip() +
  labs(x="Name", title="Top 10 Attack Pokémon")

# Top 10 Defense Pokémon
pokemon %>%
  select(Name, Defense) %>%
  arrange(desc(Defense)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(Name, Defense), y=Defense)) +
  geom_bar(stat="identity", fill="lightcyan", colour="black") +
  coord_flip() +
  labs(x="Name", title="Top 10 Defense Pokémon")

# Top 10 Special Attack Pokémon
pokemon %>%
  select(Name, Sp..Atk) %>%
  arrange(desc(Sp..Atk)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(Name, Sp..Atk), y=Sp..Atk)) +
  geom_bar(stat="identity", fill="lightcyan", colour="black") +
  coord_flip() +
  labs(x="Name", title="Top 10 Special Attack Pokémon")

# Top 10 Special Defense Pokémon
pokemon %>%
  select(Name, Sp..Def) %>%
  arrange(desc(Sp..Def)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(Name, Sp..Def), y=Sp..Def)) +
  geom_bar(stat="identity", fill="lightcyan", colour="black") +
  coord_flip() +
  labs(x="Name", title="Top 10 Special Defense Pokémon")

# Top 10 Speed Pokémon
pokemon %>%
  select(Name, Speed) %>%
  arrange(desc(Speed)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(Name, Speed), y=Speed)) +
  geom_bar(stat="identity", fill="lightcyan", colour="black") +
  coord_flip() +
  labs(x="Name", title="Top 10 Speed Pokémon")

# Top 10 Total Pokémon
pokemon %>%
  select(Name, Total) %>%
  arrange(desc(Total)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(Name, Total), y=Total)) +
  geom_bar(stat="identity", fill="lightcyan", colour="black") +
  coord_flip() +
  labs(x="Name", title="Top 10 Total Pokémon")
```

******
## Generations 
******

How many Pokémon are there per generation?

```{r}
# Number of Pokémon per generation
ggplot(pokemon, aes(x=Generation)) + 
  geom_bar(fill="lavender", colour="black") +
  labs(x="Generation", y="Number of Pokémon") 
```

The most number of Pokémon are from the first generation. Can we specify how many Pokémon of each primary type are there per generation? 

```{r}
# Number of Pokémon of each primary type per generation
ggplot(pokemon, aes(x=Type.1, fill=Generation)) + 
  geom_bar()+
  labs(x="Generation", y="Number of Pokémon") +
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

And the number of Pokémon for every type combination per generation? For example, how many Bug/Poison Pokémon are there in the second generation? It seems very complicated to calculate, but we can use the function `ftable()`.

```{r}
# The ftable is too large! We only show the type combinations with Bug as a primary type
 head(ftable(pokemon[, c("Type.1", "Type.2", "Generation")]), n=19)
```

******
## Legendary 
******

How many Legendary Pokémon are there in relation to all of the Pokémon in our data set?

```{r}
# Counts of Legendary Pokémon
table(pokemon$Legendary)
```

Only 65 Pokémon (8.8%) are Legendary! Which generation has more Legendary Pokémon? And less? Let's check it!

```{r}
# Number of Legendary Pokémon per generation
ggplot(pokemon, aes(x=Generation, fill=Legendary)) + 
  geom_bar(position="dodge") +
  labs(x="Generation", y="Number of Pokémon") 
```

******
# Summary 
******

It has been a pleasure to make this kernel! I have learned a lot, specially about data manipulation and visualization through the tools dplyr and ggplot2 from the tidyverse package. It might be interesting to re-do the analysis without considering Mega Evolutions and Legendary Pokémon

******
# References
******

Hadley Wickham (2017). tidyverse: Easily Install and Load the 'Tidyverse'. R package version 1.2.1.
https://CRAN.R-project.org/package=tidyverse

Baptiste Auguie (2017). gridExtra: Miscellaneous Functions for "Grid" Graphics. R package version 2.3.
https://CRAN.R-project.org/package=gridExtra
