```{r include = FALSE}
# This is a code block
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function


library(data.table)
library(ggplot2)
library(caret)
library(dplyr)
library(DT)
library(highcharter)
library(treemap)
```




```{r include = FALSE}
train <- fread("../input/train.csv")
test  <- fread("../input/test.csv")

dim(train)



train <- train[, timestamp :=as.Date(timestamp, format ="%Y-%m-%d")]


unique(train[, max_floor])
```


## Price Doc Densities {.tabset}


### Price_doc

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x  =  price_doc))  +geom_density() +
ggtitle("Price_doc density plot")

```


### Log price doc

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}

ggplot(train, aes(x =  log(price_doc)))  + geom_density()

```


### Floor

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x  = floor))  + geom_density() +
ggtitle("Floor density") 

```






### Product Type vs Price_doc
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x = price_doc, colour = factor(product_type), 
                  fill = factor(product_type), alpha = .3))  +  geom_density() +
                  ggtitle("Price Doc density/ Product Type")

```
### Product TYpe vs log(price_doc)

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x =  log(price_doc), colour = factor(product_type), 
                  fill = factor(product_type), alpha = .3))  + geom_density() +
                  ggtitle("Log Price Doc density / Product Type")
```

###

## Factor Variable plots {.tabset}

### thermal_power_plant_raion 
```{r warning = FALSE, echo = FALSE, message= FALSE, cache = FALSE}

ftable(train[, thermal_power_plant_raion])
prop.table(ftable(train[, thermal_power_plant_raion]))
```

```{r warning = FALSE, echo= FALSE, message = FALSE, cache = FALSE}

ggplot(train, aes(x = price_doc, 
                  fill = thermal_power_plant_raion, 
                  colour = thermal_power_plant_raion, 
                  alpha = .2)) + geom_density() +
                  ggtitle("Thermal Power Plan Raion")



ggplot(train, aes(x = log(price_doc), 
                  fill = thermal_power_plant_raion, 
                  colour = thermal_power_plant_raion, 
                  alpha = .2)) + geom_density()

```


### incineration_raion

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x = price_doc, 
                  fill = incineration_raion, 
                  colour = incineration_raion, 
                  alpha = .2)) + geom_density()


ggplot(train, aes(x = log(price_doc), 
                  fill = incineration_raion, 
                  colour = incineration_raion, 
                  alpha = .2)) + geom_density()

```



###  oil_chemistry_raion 

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x = price_doc, 
                  fill = oil_chemistry_raion, 
                  colour = oil_chemistry_raion, 
                  alpha = .2)) + geom_density()


ggplot(train, aes(x = log(price_doc), 
                  fill = oil_chemistry_raion, 
                  colour = oil_chemistry_raion, 
                  alpha = .2)) + geom_density()


```

### radiation_raion 

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x = price_doc, 
                  fill = radiation_raion, 
                  colour = radiation_raion, 
                  alpha = .2)) + geom_density()


ggplot(train, aes(x = log(price_doc), 
                  fill = radiation_raion, 
                  colour = radiation_raion, 
                  alpha = .2)) + geom_density()


```

###  railroad_terminal_raion 



```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train, aes(x = price_doc, 
                  fill = railroad_terminal_raion, 
                  colour = railroad_terminal_raion, 
                  alpha = .2)) + geom_density()


ggplot(train, aes(x = log(price_doc), 
                  fill = railroad_terminal_raion, 
                  colour = railroad_terminal_raion, 
                  alpha = .2)) + geom_density()
```

###


## Little data preprocessing 

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}

ind_num <-  sapply(train, is.character)



names_char <- names(train)[which(ind_num == 1)]



for (f in names_char) {
  if (class(train[[f]])=="character") {
    levels <- unique(c(train[[f]]))
    train[[f]] <- as.numeric(factor(train[[f]], levels=levels))
  }
}






for (j in names(train))
  set(train, which(is.na(train[[j]])), j, 0)

```

##  Timestamp Plots {.tabset}


### Simple statistics
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
# Count by timestamp 


train_count <- train %>% 
                select(timestamp, price_doc) %>%
                group_by(timestamp, price_doc) %>%
                summarize(count = n(), 
                          mean_price = mean(price_doc), 
                          median_price = median(price_doc), 
                          sd_price     = sd(price_doc))

datatable(train_count)
```
### Frequency by Timestamp

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(train_count, aes(x = timestamp, y = count  , 
                        fill=  "red", colour = "red"))  + geom_point()
                        
```

### Mean Price  
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}

ggplot(train_count, aes(x = timestamp, y = mean_price)) + 
    geom_smooth() +
    ggtitle("Mean Price/ Timestamp")
```


### Median Price

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
  ggplot(train_count, aes( x = timestamp, y = median_price))  +
  geom_smooth()  +
  ggtitle("Median Price / Timestamp")

```


###




Extract month from timestamp

```{r include = FALSE, warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
#Little processing
train  <- train[, mon_date := format(timestamp, "%m")]
train <- train[,  yearmon := format(timestamp, "%Y %m")]

```

##simple statistics , grouped by month {.tabset}


### Data Table stats
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
yearmon_group <- train %>% 
                 select(yearmon, price_doc) %>%
                 group_by(yearmon) %>%
                 summarize(mean_price  = mean(price_doc), median_price = median(price_doc), 
                           sd_price = sd(price_doc))

datatable(yearmon_group)
```



### Mean Price 
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(yearmon_group, aes(x =  yearmon, y = mean_price, 
                          fill= "red", colour = "red", 
                          ))  + geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### All at once
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(yearmon_group, aes(x = yearmon
                          ))  + geom_point(aes(y = mean_price, colour = "blue")) + 
  geom_point(aes( y = median_price, colour = "yellow")) +
  geom_point(aes(y = sd_price, colour = "black")) +
  geom_smooth(aes(y = mean_price, colour  = "blue")) +
  geom_smooth(aes( y = median_price, colour = "red")) +
  geom_smooth(aes(y = sd_price, colour = "black")) +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Bocplot with geom_jitter
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}

ggplot(train, aes(x = yearmon, y = price_doc, 
                  colour = "blue"))  +
  geom_boxplot(colour = "blue", outlier.alpha = 0.0) +
  geom_jitter(alpha = .05, colour = "red") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
###


## Highcharter Plots {.tabset}

loading time consuming :( 


```{r include = FALSE, warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
library(dplyr)
train_count_stamp <- train %>%
                     select(timestamp, price_doc) %>%
                     group_by(timestamp)  %>%
                     summarize(count = n(), 
                               mean_price = mean(price_doc), 
                               median_price = median(price_doc), 
                               sd_price = sd(price_doc))


```
### Highcharter

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}


highchart(type = "stock")    %>%
  hc_add_series_times_values(train_count_stamp$timestamp, 
                             train_count_stamp$mean_price, name = "Mean Price") %>%
  hc_add_series_times_values(train_count_stamp$timestamp, 
                             train_count_stamp$median_price, name = "median_price") %>%
  hc_add_series_times_values(train_count_stamp$timestamp, 
                             train_count_stamp$sd_price, name = "Standard Deviation of price") %>%
    hc_add_theme(hc_theme_sandsignika()) 
  
```

### Data Table
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
train_count_yearmon <- train %>%
  select(yearmon, price_doc) %>%
  group_by(yearmon)  %>%
  summarize(count = n(), 
            mean_price = mean(price_doc), 
            median_price = median(price_doc), 
            sd_price = sd(price_doc))
datatable(train_count_yearmon)


```

### Highcharter,  few stats on one plot
```{r  warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}

highchart() %>% 
  hc_xAxis(categories = train_count_yearmon$yearmon) %>%
  hc_add_series(train_count_yearmon$mean_price, type = "area", name = "Mean Price") %>%
  hc_add_series(train_count_yearmon$median_price, type = "area", name = "Median Price") %>%
  hc_add_series(train_count_yearmon$sd_price, type = "area", name = "Price Standard Deviation") %>% 
  hc_add_theme(hc_theme_sandsignika()) 

```

###


## Sub Area tree map {.tabset} 


### Count by sub_area

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}

train_desc <- fread("../input/train.csv")


train_desc_sub_area<- train_desc %>%
                      select(sub_area) %>%
                      group_by(sub_area) %>%
                    summarize(count = n())
datatable(train_desc_sub_area)
```

### Tree map plot
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
library("treemap")
library("viridisLite")

tm <- treemap(train_desc_sub_area, index = c("sub_area"),
                 vSize = "count"
                 , palette = viridis(6))
```


### Interactive Tree map plot

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
hc_tm <- highchart(height = 800) %>% 
  hc_add_series_treemap(tm, allowDrillToNode = TRUE,
                        layoutAlgorithm = "squarified",
                        name = "tmdata") %>% 
  hc_title(text = "Count by Sub Area") %>% 
  hc_tooltip(pointFormat = "<b>{point.name}</b>:<br>
             area: {point.value:,.0f}<br>
             count: {point.valuecolor:,.0f}")
             
 hc_tm            
```



20170502 - end. 

# Differences in test dataset 

## Sub Area tree map TEST SET {.tabset} 


### Count by sub_area

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}

test_desc <- fread("../input/test.csv")


test_desc_sub_area<- test_desc %>%
                      select(sub_area) %>%
                      group_by(sub_area) %>%
                    summarize(count = n())
datatable(test_desc_sub_area)
```

### Tree map plot - test dataset 
```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}


tm <- treemap(test_desc_sub_area, index = c("sub_area"),
                 vSize = "count"
                 , palette = viridis(6))
```


### Interactive Tree map plot

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
hc_tm <- highchart(height = 800) %>% 
  hc_add_series_treemap(tm, allowDrillToNode = TRUE,
                        layoutAlgorithm = "squarified",
                        name = "tmdata") %>% 
  hc_title(text = "Count by Sub Area") %>% 
  hc_tooltip(pointFormat = "<b>{point.name}</b>:<br>
             area: {point.value:,.0f}<br>
             count: {point.valuecolor:,.0f}")
             
 hc_tm            
```



### Simple statistics

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
# Count by timestamp 


test_count <- test_desc %>% 
                select(timestamp) %>%
                group_by(timestamp) %>%
                summarize(count = n())

datatable(test_count)
```
### Frequency by Timestamp

```{r warning = FALSE, echo = FALSE, message = FALSE, cache = FALSE}
ggplot(test_count, aes(x = timestamp, y = count  , 
                        fill=  "red", colour = "red"))  + geom_point()
                        
```

Update 20170505
does not make sense to look deeper in test dataset.
Todo: independent kernel with lasso, l1, penalized and trees. 
end.